---
title: "p8105_hw6_am6594"
author: "Alice Mao"
date: "2024-12-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load necessary libraries
library(tidyverse)
library(broom)
library(purrr)
library(modelr)
library(crossval)
library(scales)
library(rnoaa)

```

# Problem 1

```{r}
# Download data
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

# Preview data
head(weather_df)

```
```{r}
# Fit the linear regression model
lm_model <- lm(tmax ~ tmin, data = weather_df)
summary(lm_model)

# Extract R^2
r_squared <- glance(lm_model)$r.squared
r_squared

```

```{r}
# Define the bootstrap function
set.seed(123)
bootstrap_samples <- weather_df %>%
  bootstrap(5000)

# Perform bootstrap analysis
bootstrap_results <- bootstrap_samples %>%
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    r_squared = map_dbl(models, ~glance(.x)$r.squared),
    log_beta = map(models, ~tidy(.x)) %>%
      map_dbl(~log(.[1, "estimate"] * .[2, "estimate"]))
  )

# Extract R^2 and log(beta0 * beta1)
r_squared_bootstrap <- bootstrap_results$r_squared
log_beta_bootstrap <- bootstrap_results$log_beta

# Calculate 95% Confidence Intervals
r_squared_ci <- quantile(r_squared_bootstrap, c(0.025, 0.975))
log_beta_ci <- quantile(log_beta_bootstrap, c(0.025, 0.975))
```


# Problem 2

```{r}
# Load the dataset
homicides <- read_csv("data/homicide-data.csv")
```

```{r}
# Create city_state variable
homicides <- homicides %>%
  mutate(city_state = paste(city, state, sep = ", "))

# Create binary variable indicating whether the homicide is solved
homicides <- homicides %>%
  mutate(solved = if_else(disposition == "Closed by arrest", 1, 0))

# Omit specified cities
omit_cities <- c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
homicides <- homicides %>%
  filter(!city_state %in% omit_cities)

# Limit analysis to cases where victim_race is "White" or "Black"
homicides <- homicides %>%
  filter(victim_race %in% c("White", "Black"))

# Ensure victim_age is numeric and remove rows with missing victim_age
homicides <- homicides %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(!is.na(victim_age))

```

```{r}
# Filter data for Baltimore, MD
baltimore_data <- homicides %>%
  filter(city_state == "Baltimore, MD")

# Fit logistic regression model
baltimore_glm <- glm(solved ~ victim_age + victim_sex + victim_race, data = baltimore_data, family = binomial)

# Save output as an R object and tidy the results
baltimore_tidy <- tidy(baltimore_glm, conf.int = TRUE, exponentiate = TRUE)

# Obtain estimate and CI for adjusted OR comparing male to female victims
male_vs_female_or <- baltimore_tidy %>%
  filter(term == "victim_sexMale") %>%
  select(term, estimate, conf.low, conf.high)

# Display the result
male_vs_female_or
```



```{r}
# Filter out cities where victim_sex or victim_race has only one level
city_sex_race_counts <- homicides %>%
  group_by(city_state) %>%
  summarize(
    sex_levels = n_distinct(victim_sex),
    race_levels = n_distinct(victim_race)
  )

valid_cities <- city_sex_race_counts %>%
  filter(sex_levels > 1, race_levels > 1) %>%
  pull(city_state)

homicides_filtered <- homicides %>%
  filter(city_state %in% valid_cities)

# Run glm for each city with error handling
city_models <- homicides_filtered %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ safely(glm)(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial))
  )

# Extract the results, handling errors
city_models <- city_models %>%
  mutate(model_result = map(model, "result"), model_error = map(model, "error"),
         tidied = map(model_result, ~ if(!is.null(.x)) tidy(.x, conf.int = TRUE, exponentiate = TRUE) else NULL))

# Extract adjusted ORs and CIs for victim_sexMale
or_results <- city_models %>%
  unnest(tidied) %>%
  filter(term == "victim_sexMale") %>%
  select(city_state, estimate, conf.low, conf.high)

```

```{r}
# Organize cities according to estimated OR
or_results <- or_results %>%
  arrange(estimate) %>%
  mutate(city_state = factor(city_state, levels = city_state))

# Create the plot
ggplot(or_results, aes(x = estimate, y = city_state)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  xlab("Adjusted Odds Ratio (Male vs Female Victims)") +
  ylab("City") +
  ggtitle("Adjusted Odds Ratios for Solving Homicides by City") +
  theme_minimal()

```

This plot shows adjusted odds ratios (ORs) for solving homicides by city, comparing male versus female victims. The x-axis represents the odds ratio, with a red reference line at OR = 1, indicating no difference between male and female victims in terms of case resolution. Values above 1 suggest that cases involving male victims are more likely to be solved, while values below 1 indicate a higher likelihood for cases involving female victims. While the majority of cities have ORs near 1, indicating similar likelihoods of solving cases for both genders, some cities show substantial deviations. Notably, the wide confidence intervals for some cities indicate a lack of precision, possibly due to smaller sample sizes or greater variability in the data. These discrepancies suggest systemic or regional differences in investigative priorities or resources, warranting further exploration of local practices and socio-cultural factors influencing case resolutions. The plot underscores the importance of understanding such disparities to ensure equity in homicide investigations across gender and location.

# Problem 3


```{r}
# Load the dataset
birthweight <- read_csv("data/birthweight.csv")

# Briefly check the dataset
structure(birthweight)
```

```{r}
# Convert appropriate variables to factors
birthweight <- birthweight %>%
  mutate(
    babysex = factor(babysex, labels = c("Male", "Female")),
    frace = factor(frace),
    malform = factor(malform, labels = c("Absent", "Present")),
    mrace = factor(mrace)
  )

# Check for missing data
missing_data <- colSums(is.na(birthweight))

# Summarize key variables
summary(birthweight)

# Visualize birthweight distribution
ggplot(birthweight, aes(x = bwt)) +
  geom_histogram(binwidth = 100, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Birthweights", x = "Birthweight (grams)", y = "Count")

```


```{r}
# Hypothesized model
model_hypothesis <- lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + ppbmi + smoken, data = birthweight)

# Add predictions and residuals
birthweight <- birthweight %>%
  add_predictions(model_hypothesis, var = "fitted") %>%
  add_residuals(model_hypothesis, var = "residuals")

# Plot residuals vs. fitted values
ggplot(birthweight, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals")

```

```{r}
set.seed(123)

# Create cross-validation folds
cv_data <- crossv_mc(birthweight, n = 10)

# Fit and evaluate models
cv_results <- cv_data %>%
  mutate(
    model_hypothesis = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + ppbmi + smoken, data = as.data.frame(.))),
    model_main_effect = map(train, ~lm(bwt ~ blength + gaweeks, data = as.data.frame(.))),
    model_interact = map(train, ~lm(bwt ~ bhead * blength * babysex, data = as.data.frame(.))),
    
    # Calculate RMSE for test sets
    rmse_hypothesis = map2_dbl(model_hypothesis, test, ~sqrt(mean((predict(.x, as.data.frame(.y)) - as.data.frame(.y)$bwt)^2))),
    rmse_main_effect = map2_dbl(model_main_effect, test, ~sqrt(mean((predict(.x, as.data.frame(.y)) - as.data.frame(.y)$bwt)^2))),
    rmse_interact = map2_dbl(model_interact, test, ~sqrt(mean((predict(.x, as.data.frame(.y)) - as.data.frame(.y)$bwt)^2)))
  )

# Gather RMSE results
rmse_results <- cv_results %>%
  select(rmse_hypothesis, rmse_main_effect, rmse_interact) %>%
  pivot_longer(cols = everything(), names_to = "model", values_to = "rmse") %>%
  mutate(model = recode(model,
                        rmse_hypothesis = "Hypothesis",
                        rmse_main_effect = "Main Effect",
                        rmse_interact = "Interactions"))

```


```{r}
# Plot RMSE distributions
ggplot(rmse_results, aes(x = model, y = rmse)) +
  geom_violin() +
  labs(title = "Cross-Validation RMSE Distributions",
       x = "Model",
       y = "RMSE") +
  theme_minimal()

# Calculate average RMSE for each model
rmse_summary <- rmse_results %>%
  group_by(model) %>%
  summarize(mean_rmse = mean(rmse), .groups = "drop")

# Display RMSE summary
rmse_summary
```

Based on the results, the hypothesis model is the most accurate and consistent predictor of birthweight. In contrast, the main effect model is the least accurate, likely due to the exclusion of important variables and interaction terms. Although the interactions model is more complex than the hypothesis model, it does not significantly improve performance and introduces greater variability.


